\documentclass{notes}


% ********** uncomment lines 5,6,7 for a dark pdf
% \usepackage{xcolor}
% \pagecolor[rgb]{0,0,0} %black
% \color[rgb]{0.5,0.5,0.5} %grey
%*******************



\begin{document}

%CHANGE THE INFO THAT APPEARS ON THE TITLE PAGE AND THE HEADERS
\instructor{Instructor: Valentino \textsc{Tosatti}\par}
\courseid{MATH 466}
\author{Alexander Kroitor}
\version{Fall 2020}
\university{McGill University}

\begin{titlepage}
%CHANGE THE TITLE OF THE PAPER
    \title{Complex Analysis}
    \maketitle
\end{titlepage}



\setcounter{lecnum}{0}
\setcounter{page}{1}
\tableofcontents
%\thispagestyle{empty}

\begin{note}[From Scribe]
This is a set of notes for MATH466 at McGill taught by Valentino Tosatti in the Fall of 2020. For any questions, comments, corrections, requests, complaints, or insults please feel free to e-mail alexander.kroitor@mail.mcgill.ca.\\
 
 I will happily provide the most recent set of notes I have made and the source .tex files. These notes are intended to follow the lectures exactly. Some of the less compelling examples have been excluded.
\end{note}

\begin{note}[To Scribe and Interested Readers]
All lectures, especially 6 and 7, deserve a thorough review for readability. The section on Ascoli in chapter 7 is confused, and needs the insight from lecture 8 included. All lectures need to have appropriate section/subsection markers added to them, especially invisible ones for important results. The format and style change as I go and improve at latex, so earlier chapters should be revisited to standardize them.\\

I need to go back and change up the old notation to be perfectly consistent -- an example is the old $\HH = \{ x \in \C \mid \Im(x) > 0\}$ format versus the more graphically and syntactically correct $\HH = \set{ x \in \C \mid \Im(x) > 0 }$ format.\\

Note further that this \LaTeX\ code is meant to be compiled in pdfLaTeX, and will not work in other compilers (to my knowledge).
\end{note}




% \input{lectures/lec01}
% \input{lectures/lec02}
% \input{lectures/lec03}
% \input{lectures/lec04}
% \input{lectures/lec05}
\input{lectures/lec06}
% \input{lectures/lec07}
% \input{lectures/lec08}
% \input{lectures/lec09}
% \input{lectures/lec10}
% \input{lectures/lec11}
% \input{lectures/lec12}
% \input{lectures/lec13}
% \input{lectures/lec14}
% \input{lectures/lec15}
% \input{lectures/lec16}
% \input{lectures/lec17}
% \input{lectures/lec18}
% \input{lectures/lec19}
% \input{lectures/lec20}
% \input{lectures/lec21}
% \input{lectures/lec22}
% \input{lectures/lec23}
% \input{lectures/lec24}
% \input{lectures/lec25}
% \input{lectures/lec26}

\end{document}